Constraint Programming - Inside Gecode

### 2014-05-03 paper.tex
correction
constraint propagation

	\subsection{Constraint Types}
	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like [equation \ref{eq:linearConstraint}]. Which statements are possible to make varies than very much from the programming language in which the problem is implemented and the used toolset. Typical constraints would be
	\begin{itemize}
		\item Equalities: $x = y + 2$, $x^2 + y^2 = 1$, \dots
		\item Inequalities: $x \ne y$, \dots
		\item Relations: $x < 2 y$, $e^x < 4$, \dots
	\end{itemize}
	In Gecode for example it is very easy to create for example a nonlinear constraint based on the integer variables a, b, c and d \cite[p. 120]{programmingGecode}:
	\begin{lstlisting}
rel(home, a+b*(c+d) == 0);
	\end{lstlisting}
	This single statement in C++ implements the constraint
	\begin{equation}
		a + b \cdot (c + d) == 0
	\end{equation}
	what is a quite handy way to do it.

@@

	\subsection{Constraint Propagation}

>>>>

	\subsection{Constraint Types}
	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like [equation \ref{eq:linearConstraint}]. Which statements are possible to make varies than very much from the programming language in which the problem is implemented and the used toolset. Typical constraints would be
	\begin{itemize}
		\item Equalities: $x = y + 2$, $x^2 + y^2 = 1$, \dots
		\item Inequalities: $x \ne y$, \dots
		\item Relations: $x < 2 y$, $e^x < 4$, \dots
	\end{itemize}
	In Gecode for example it is very easy to create for example a nonlinear constraint based on the integer variables a, b, c and d \cite[p. 120]{programmingGecode}:
	\begin{lstlisting}
rel(home, a+b*(c+d) == 0);
	\end{lstlisting}
	This single statement in C++ implements the constraint
	\begin{equation}
		a + b \cdot (c + d) == 0
	\end{equation}
	which is a quite handy way to do it.

@@

	\subsection{Constraint Propagation}
	Constraint propagation is a key concept in constraint programming as it allows to speed-up the search significantly. The term propagation refers in this area to the reduction of domains of variables which are infeasible, considering other constraints and affected variable domains. As the size of the variable domain is related to possibilities which have to be checked by a search, the constraint propagation is able to prune infeasible subtrees. Therefore the whole search process is accelerated.

	The actual implementation of constraint propagation differs again from one toolset to another, but the basic concepts are for sure the same. As the topic of this paper is the implementation in Gecode I will concentrate on this one.

	Constraints in Gecode are always evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the subtrees. The problem here is to decide in which order the constraints should be considered, and it may even be necessary to evaluate some constraints more often. A short example will explain this very well \cite[p. 575]{handbookCP}: Let there be two constraints 
	\begin{equation}
	c_1: y - x^2 = 0
	\end{equation}
	\begin{equation}
	c_2: y - x - 1 = 0
	\end{equation}
	and let the variable domains be $x, y \in [-4, 4]$. Starting with the evaluation of constraint one follows a chain of reductions of the variable domains, based only on constraint propagation:
	\begin{equation}
	\begin{split}
	c_1:\ &y \in [-4, 4] \rightarrow y \in [0, 4] \\
	c_1:\ &x \in [-4, 4] \rightarrow x \in [-2, 2] \\
	c_2:\ &y \in [0, 4] \rightarrow y \in [0, 3] \\
	c_2:\ &x \in [-2, 2] \rightarrow x \in [-1, 2] \\
	c_1:\ &x \in [-1, 2] \rightarrow x \in [-1, 1.73 \dots] \\
	&\dots
	\end{split}
	\end{equation}
	Actually this behaviour is even very nice to have as it narrows the down the search space very efficient.

	To implement this repeated evaluation of constraints in Gecode all constraints subscribe to the domains of the variables which are part of them. With this method the constraints get noticed whenever a variable domain is reduced and they can signal the system that they need to be evaluated once again.

	The next difficult decision is which constraint should evaluated first. In the example above the constraints were quite simple and therefore it won't make a difference to evaluate first $c_1$ or $c_2$, the search will take nearly the same time. Unfortunately not all constraints are that easy to evaluate and for this case Gecode sets for every constraint a certain cost \cite[p. 275]{programmingGecode}. Based on this cost the constraints are sorted into buckets which are then iterated from the constraints with lower costs to the ones with higher costs. These means, whenever a variable domain is changed the first try is to use constraints with the lowest possible cost and only if none of them needs to be evaluated constraints with higher costs are considered.
